---
title: "linear model from Web"
author: "Mairani G"
date: "Summer 2024"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

# Methods
All variables are numerical, so we use num~num helper file for all 3 parts.  

```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS

### descriptive results -- scatterplot

```{r}
plot(SIMS~ARM,data=data)
```

The plot on the SIMS score based on ARM strength shows that as arm strength got better, the SIMS score became higher.  


### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```
The correlation shows that a positive relationship exists between both variables. Confirming that as arm strength increases, so does the SIMS score.  

### Inferential  (Build model.1)

```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```

#### scatterplot with model fit

```{r}
plot(SIMS~ARM,data=data)
abline(model.1,col='red')
```

The p-value is < 2.2e-16. In its scientific notation, this is 0.00000000000000016. It is really close to zero and smaller than 0.05.   

#### Predict at target point

```{r}
newdata = data.frame(ARM = 88, GRIP=94)
predict(model.1, newdata, interval='pred')
```

The SIMS prediction for ARM =88 is 0.7063836, the predicting interval is (-1.73, 3.14)


## Model 2 SIM~GRIP

### Descriptive results -- scatterplot

```{r}
plot(SIMS~GRIP,data=data)
```

The scatter plot shows that as Grip strength is stronger, the SIMS score typically increases.  

### Numerical results 

```{r}
cor(SIMS~GRIP,data=data)
```
The correlation 0.6398458 shows that there is a positive relationship. The grip strength increases.   

### Inferential  (Build model.2)

```{r}
model.2 <- lm(SIMS~GRIP,data=data)
summary.lm(model.2)
```
#### scatterplot with model fit

```{r}
plot(SIMS~GRIP,data=data)
abline(model.2,col='red')
```

The p-value is < 2.2e-16, the same as the arm strength. It is 0.00000000000000016 and smaller than 0.05.  

#### predict model.2 at target point

#### now add the model fit to our plot for model.2
  


## Model 3 SIM~ARM+GRIP

### Numerical results (cor)

`
  
### Inferential  (Build 2-dimentional model.3)

#### predict model.3 at target point
  

## Comparing nested models ANOVA Test

### Model.1 vs Model.3
### Model.2 vs Model.3

## Informally compare Model.1 with model.2

