---
title: "linear model from Web"
author: "Mairani G"
date: "Summer 2024"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

# Methods
All variables are numerical, so we use num~num helper file for all 3 parts.  

```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS

### descriptive results -- scatterplot

```{r}
plot(SIMS~ARM,data=data)
```

The plot on the SIMS score based on ARM strength shows that as arm strength got better, the SIMS score became higher.  


### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```
The correlation shows that a positive relationship exists between both variables. Confirming that as arm strength increases, so does the SIMS score.  

### Inferential  (Build model.1)

```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```

#### scatterplot with model fit

```{r}
plot(SIMS~ARM,data=data)
abline(model.1,col='red')
```

The p-value is < 2.2e-16. In its scientific notation, this is 0.00000000000000016. It is really close to zero and smaller than 0.05. The model fit shows an upwards trend.   

#### Predict at target point

```{r}
newdata = data.frame(ARM = 88, GRIP=94)
predict(model.1, newdata, interval='pred')
```

The SIMS prediction for ARM =88 is 0.7063836, the predicting interval is (-1.73, 3.14)


## Model 2 SIM~GRIP

### Descriptive results -- scatterplot

```{r}
plot(SIMS~GRIP,data=data)
```

The scatter plot shows that as Grip strength is stronger, the SIMS score typically increases.  

### Numerical results 

```{r}
cor(SIMS~GRIP,data=data)
```
The correlation 0.6398458 shows that there is a positive relationship. The grip strength increases.   

### Inferential  (Build model.2)

```{r}
model.2 <- lm(SIMS~GRIP,data=data)
summary.lm(model.2)
```
#### scatterplot with model fit

```{r}
plot(SIMS~GRIP,data=data)
abline(model.2,col='red')
```

The p-value is < 2.2e-16, the same as the arm strength. It is 0.00000000000000016 and smaller than 0.05. The model fit shows an upwards trend.   

#### predict model.2 at target point
```{r}
newdata = data.frame(ARM=88, GRIP=94)
predict(model.2,newdata,interval='pred')
```
The SIMS prediction for ARM = 88 is 0.5361543 , the predicting interval is (-3.11,2.04)  


## Model 3 SIM~ARM+GRIP

### numerical results 

### numerical results (cor)
```{r}
df=data.frame(SIMS=data$SIMS,GRIP=data$GRIP,ARM=data$ARM)
cor(df)
```

Correlation matrix shows that SIMS correlates with GRIP and ARM, also that GRIP and ARM also corrlate with each other.  

### Inferential  (Build 2-dimentional model.3)

```{r}
model.3=lm(SIMS~ARM+GRIP,data=data)
summary(model.3)
```

The p-value from the F-statistic is 0.00000000000000022 and it shows that the relationship between the sims score and arm and grip strengths is not coincidental. It is less than 0.05 and does fit the model.

p-values from the model parameters: intercept is < 2e-16, coefficient of ARM is 1.50e-09 and coefficient of GRIP is 4.99e-06.  

#### predict model.3 at target point
  
```{r}
newdata=data.frame(GRIP=94, ARM= 88)
predSIMS= predict(model.3, newdata, interval= 'pred')
print(predSIMS)
```
The predicted SIMS score for GRIP=94, ARM=88 is 0.150.The predicting interval is (-2.13,2.43) 

## Comparing nested models ANOVA Test

### Model.1 vs Model.3

```{r}
anova(model.1, model.3)
```

The anova shows the p-value is 4.994e-06 (0.000004994), it is less than 0.05, so we reject the null hypothesis and it is shown that there is a difference in the models.  

Model 2 : SIMS~ARM+GRIP shows a smaller residual sum of squared errors, therefore it is a better fit than model 1:SIMS~ARM. The rss proves this as it is 188 in model 2 and 218 in model 1.  

### Model.2 vs Model.3

```{r}
anova(model.2, model.3)
```

## Informally compare Model.1 with model.2

